---
title: "p8105_hw3_jer2243"
author: "Juliet Rowe"
date: "2023-10-09"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

# Problem 1


## Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

## Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```



# Problem 2

## Load packages and BRFSS data
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(p8105.datasets)
data("brfss_smart2010")
```


## Clean BRFSS data
```{r}
levels <- c("Poor", "Fair", "Good", "Very good", "Excellent")
values <- c(1, 2, 3, 4, 5)

brfss_smart2010_tidy = brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  transform(response = factor(response, levels = levels, labels = values, ordered = TRUE))
  
```
This dataset contains `r nrow(brfss_smart2010_tidy)` rows and `r ncol(brfss_smart2010_tidy)` columns. 

## In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_2002 <- brfss_smart2010_tidy |>
  filter(year==2002)

result1 <- brfss_2002 |>
  group_by(locationabbr) |>
  summarise(unique_locationdesc = n_distinct(locationdesc)) |>
  filter(unique_locationdesc >= 7)

print(result1)

brfss_2010 <- brfss_smart2010_tidy |>
  filter(year==2010)

result2 <- brfss_2010 |>
  group_by(locationabbr) |>
  summarise(unique_locationdesc = n_distinct(locationdesc)) |>
  filter(unique_locationdesc >= 7)

print(result2)
```
In 2002, `r nrow(filter(result1, unique_locationdesc >= 7))` states were observed at 7 or more locations. These included `r pull(filter(result1, unique_locationdesc >=7), locationabbr)`. In 2010, `r nrow(filter(result2, unique_locationdesc >= 7))` states were observed at 7 or more locations. These included `r pull(filter(result2, unique_locationdesc >=7), locationabbr)`.


## Construct a dataset limited to "excellent" responses
```{r}
brfss_excellent = brfss_smart2010_tidy |>
  filter(response == "5") |>
  group_by(locationabbr, year) |>
  mutate(data_value_avg = mean(data_value, na.rm = TRUE)) |>
  select(year, locationabbr, data_value_avg)
  
```
The dataset limited to "excellent" responses consists of `r nrow(brfss_excellent)` observations.

## Spaghetti plot of the avg value over time within a state
```{r}
ggplot(brfss_excellent, aes(x=year, y=data_value_avg, color=locationabbr, group=locationabbr)) + 
  geom_line() +
  labs(x="Year", y="Average Data Value", color = "State") +
  ggtitle("Average Data Value by State over Years")
```
In general, the average data value has fluctuated over the years but appears to have decreased slightly overall. Most average data values fall between 17.5 and 27.5, with a few outliers on either end. 

## Two-panel plot
```{r}
ny_data <- brfss_smart2010_tidy |>
  filter(locationabbr == "NY" & year %in% c(2006, 2010)) |>
  separate(locationdesc, into =c("state", "county"), sep=5) |>
  select(-state) |>
  group_by(county)

ggplot(ny_data, aes(x = data_value, y = response)) +
  geom_point(aes(color=county)) +
  facet_grid(.~year) +
  ggtitle("Distribution of Data Value by Response in NY State")
```
'Data value' values were the highest for the responses "good" and "very good". Overall, data_value increased from 2006 to 2010, and the range of scores per response category also widened.


# Problem 3


## Import and clean demographic and accelerometer data
```{r}
demo_data = read_csv(file = "~/Desktop/p8105_hw3_jer2243/nhanes_covar.csv", skip=4) |>
  janitor::clean_names()

accel_data = read_csv(file = "~/Desktop/p8105_hw3_jer2243/nhanes_accel.csv") |>
  janitor::clean_names()
```


## Clean the demo dataset
```{r}
demo_data <- subset(demo_data, age>=21)
demo_data <- demo_data[complete.cases(demo_data), ]
```


## Merge demo dataset with accel dataset
```{r}
merged_data = 
  left_join(demo_data, accel_data, by = "seqn")
```
The full merged dataset with demographic and accelerometer data contains `r nrow(merged_data)` observations.

## Encode data with reasonable variable classes

```{r}
merged_data_2 = merged_data |>
  mutate(education = case_when(
    education == 1 ~ "less than HS",
    education == 2 ~ "HS equivalent",
    education == 3 ~ "more than HS",
    TRUE ~ as.character(education)),
    sex = case_when(
      sex == 1 ~ "male",
      sex == 2 ~ "female",
      TRUE ~ as.character(sex)
    )
  )
```


## Number of men and women in each education category
```{r}
merged_data_2 |>
  group_by(sex, education) |>
  summarize (count=n()) |>
  knitr::kable()
```
For both males and females, the highest number of observations was in the "more than high school" education category. The lowest count for females was in the "HS equivalent" category, and the lowest count for males was in the "less than HS" category.

## Graph the age distributions for men and women in each education category
```{r}
merged_data_2 |>
  ggplot(aes(x=age)) +
  geom_histogram(bins=10) +
  facet_wrap(sex~education) +
  labs(title = "Age distributions for men and women by education category")

```
Females in the highest education group tended to be younger, while males in the highest education group tended to be younger than 40 or above 60. Males and females in the "HS equivalent" or "less than HS" groups tended to be older or middle aged, and the distributions across sex appear fairly similar.

## Create total activity variable
```{r}
merged_data_3 = merged_data_2 |>
  mutate(
    total_activity = rowSums(across(c(min1:min1440))
  ))
```

## Create plots of total activity by age, comparing men and women, separate panels for each educ level
```{r}
ggplot(merged_data_3, aes(x = age, y = total_activity, color=sex)) +
  geom_point() +
  geom_smooth(se=FALSE) +
  facet_wrap(~ education) + 
  labs(title = "Total activity by age for men and women across education levels")
```
Overall, total activity tended to decrease as age increased for both sexes and all education groups. In the "HS equivalent" and "more than HS" groups, females tended to be more active than males, while in the "less than HS" group, males were slightly more active than females. The biggest difference in activity level between males and females was seen for those in the "HS equivalent" group between the ages of 30 and 50.

## Make a 3-panel plot of the 24-hr activity time course for each education level and use color to indicate sex
```{r}
merged_data_4 = merged_data_3 |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "activity"
    )

merged_data_4 |>
  ggplot(aes(x=minute, y=activity, color=sex)) + geom_point(alpha=0.25, size=0.1) + geom_smooth() + facet_wrap(.~education) +
  labs(title = "24-hour activity time course by education level and sex")

```
For all three panels, there was a dip in activity levels during the early morning hours, and the highest activity levels were around midday. Overall, it appears as though the "more than HS" group had slightly higher activity levels compared to the other two groups.

## Repeat plot above but without outliers where activity >50
```{r}
merged_data_4 |>
  filter(activity < 50) |>
  ggplot(aes(x=minute, y=activity, color=sex)) + geom_point(alpha=0.5, size=0.5) +
  geom_smooth(linewidth = 2) + facet_wrap(.~education) +
  labs(title = "24-hour activity time course by education level and sex")
```
Excluding the outliers where activity > 50 allows to get a slightly more useful, zoomed in view of the data. 
